---
title: "Phoenix Arizona"
output: html_notebook
---

This will be all the work for the Phoenix, Arizona data (one of the 5 selected cities used as part of this thesis work).

```{r}
library(ggmap)
library(glue)
library(here)
library(osmdata) # package for working with streets
library(sf)
library(showtext) # for custom fonts
library(spData)
library(spDataLarge)
library(terra)
library(tidyverse)
library(tmap)
library(rvest)
library(tidyverse)
library(osmdata) # package for working with streets
library(showtext) # for custom fonts
library(ggmap)
library(rvest)
library(ggplot2) # added this one myself
library(sf)
library(gapminder) # might work for image saving?
library (igraph)
library (ggraph)
library (h3r)
```

```{r}
getbb("Phoenix Arizona")
available_tags("highway")

big_streets <- getbb("Phoenix Arizona")%>%
  opq()%>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary", "motorway_link", "primary_link")) %>%
  osmdata_sf()

med_streets <- getbb("Phoenix Arizona")%>%
  opq()%>%
  add_osm_feature(key = "highway", 
                  value = c("secondary", "tertiary", "secondary_link", "tertiary_link")) %>%
  osmdata_sf()


small_streets <- getbb("Phoenix Arizona")%>%
  opq()%>%
  add_osm_feature(key = "highway", 
                  value = c("residential", "living_street",
                            "unclassified",
                            "service", "footway")) %>%
  osmdata_sf()

phoenixarizonadata <- bind_rows(big_streets$osm_lines, med_streets$osm_lines, small_streets$osm_lines)

save(phoenixarizonadata, 
     file = glue::glue(here::here(), "/data-raw/phoenixarizonadata.rda"), 
     compress = "xz")
```
Determining Urban vs Suburban Areas Using Geohashes:

```{r}
getbb("Phoenix Arizona")
getPentagons("6") #h3r package

phoenix_boundaries <- getbb("Phoenix Arizona")%>%
  opq()%>%
  add_osm_feature(key = "boundary", 
                  value = c("statistical")) %>%
  osmdata_sf()


geohashes <- h3r::_(phoenix_boundaries, res = 6)
  
  # if statistical does not work, try census


```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
