---
title: "Phoenix Arizona"
output: html_notebook
---

This will be all the work for the Phoenix, Arizona data (one of the 5 selected cities used as part of this thesis work).

Load packages:
```{r}
#library(ggmap)
library(glue) # Interpreted String Literals
library (h3)
library(here) # A Simpler Way to Find Your Files
library(osmdata) # package for working with streets
library(sf) # Simple Features for R
library(tidycensus) # Load US Census Boundary and Attribute Data as 'tidyverse' and 'sf'-Ready Data Frames
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(tigris) # Load Census TIGER/Line Shapefiles
```

## Network data

Download OSM data for Arizona [here](https://download.geofabrik.de/north-america/us/arizona.html). It will be a proto-buffer file (pbf). Save under `data-raw/r5_graph`.

## Census tracts

Obtain the census tracts for Phoenix. Note that [{tidycensus}](https://walker-data.com/tidycensus/) function `get_decenial()` allows the selection of state and county, but not metro area or city. Check the example [here](https://walker-data.com/census-r/spatial-analysis-with-us-census-data.html#identifying-geometries-within-a-metropolitan-area) to see how to select tracts in a city.

```{r}
options(tigris_use_cache = TRUE)

az_tracts <- get_decennial(geography = "tract",
                           variables = "H1_001N",
                           state = "Arizona",
                           year = 2020,
                           geometry = TRUE
)

az_city <- urban_areas(year = 2020) |> 
  filter(str_detect(NAME10, "AZ"))
```

A [core-based statistical area](https://en.wikipedia.org/wiki/Core-based_statistical_area) is a generous definition of the area of interest, since it contains an urban nucleus (say, Phoenix) and also all neighboring communities that are tightly integrated with it.
```{r}
phoenix_cbsa <- core_based_statistical_areas(year = 2020) |>
  filter(str_detect(NAME, "Phoenix"))
```

A much stricter definition of the area of interest would result from using the city boundaries, which can be obtained using `urban_areas()`:
```{r}
phoenix_city <- urban_areas(year = 2020) |> 
  filter(str_detect(NAME10, "Phoenix--Mesa"))
```

Plot:
```{r}
ggplot() +
  geom_sf(data = az_tracts,
          fill = "lightgrey") +
  geom_sf(data = phoenix_cbsa,
          color = "red",
          fill = NA) +
  geom_sf(data = az_city,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

Perhaps select all urban tracts within Phoenix CBSA? First select all urban areas in Phoenix CBSA:
```{r}
phoenix_urban <- st_filter(az_city, phoenix_cbsa)
```

Plot:
```{r}
ggplot() +
  geom_sf(data = phoenix_cbsa,
          fill = "lightgray") +
  geom_sf(data = phoenix_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

Notice that there is a big urban area that is mostly outside of the Phoenix CBSA (Tucson). This will need to be removed. 

```{r}
phoenix_urban <- phoenix_urban |>
  filter(!str_detect(NAME10, "Tucson"))
```

Verify:
```{r}
ggplot() +
  geom_sf(data = phoenix_cbsa,
          fill = "lightgray") +
  geom_sf(data = phoenix_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

Obtain the tracts:
```{r}
phoenix_tracts <- az_tracts |>
  st_intersection(phoenix_urban)
```

Plot:
```{r}
ggplot() +
  geom_sf(data = phoenix_cbsa,
          fill = "lightgray") +
  geom_sf(data = phoenix_tracts,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

## Geohashes

Use the centroid of Phoenix CBSA to obtain h3 polygons:
```{r}
coords <- st_centroid(phoenix_cbsa) |>
  st_coordinates()
```


Retrieve h3 geohashing at a fairly high resolution to identify Hamilton; the coordinates are for the City of Hamilton:
```{r}
phoenix_h3 <- geo_to_h3(c(coords[2], coords[1]),
                        res = 3) |>
  k_ring(radius = 1)
```

Plot the geohash retrieved in this way:
```{r}
ggplot() + 
  geom_sf(data = h3_to_geo_boundary_sf(phoenix_h3),
          aes(fill = h3_index)) +
  geom_sf(data = phoenix_cbsa,
          color = "red",
          fill = NA) +
  geom_sf(data = phoenix_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

Keep only the h3 indices that overlap with urban Phoenix:
```{r}
phoenix_h3 = c("8348e9fffffffff", "8348e8fffffffff", "8348ebfffffffff", "8329b6fffffffff")
```

Plot kept indices:
```{r}
ggplot() + 
  geom_sf(data = h3_to_geo_boundary_sf(phoenix_h3),
          aes(fill = h3_index)) +
  geom_sf(data = phoenix_cbsa,
          color = "red",
          fill = NA) +
  geom_sf(data = phoenix_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

Obtain children at a higher resolution:
```{r}
phoenix_h3 <- phoenix_h3 |> 
  purrr::map(\(x) h3_to_children(x, res = 8)) |>
  unlist()
```

Choose only hexagons that touch urban Phoenix:
```{r}
phoenix_h3 <- h3_to_geo_boundary_sf(phoenix_h3) |>
  st_transform(crs = st_crs(phoenix_urban)) |>
  st_intersection(phoenix_urban)

phoenix_h3 <- phoenix_h3 |> 
  pull(h3_index)
```

Plot:
```{r}
ggplot() + 
  geom_sf(data = h3_to_geo_boundary_sf(phoenix_h3)) +
  geom_sf(data = phoenix_cbsa,
          color = "red",
          fill = NA) +
  geom_sf(data = phoenix_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

Currently this is resolution 8:
```{r}
h3_get_resolution(phoenix_h3)
```

Obtain children at resolution 13 (average area is 43.87 $m^2$):
```{r}
phoenix_h3 <- phoenix_h3 |> 
  purrr::map(\(x) h3_to_children(x, res = 13)) |>
  unlist()
```

Check the size of an h3 hexagon:
```{r}
sample_h3 <- h3_to_geo_boundary_sf(phoenix_h3[1]) 

sample_h3 |> 
  mutate(area = st_area(geometry))
```

Information about housing characteristics. Explore the tables [here](https://data.census.gov/advanced?q=Housing) to get a sense of the variables that might be interesting to date the housing stock. For example, a search within the topic of "housing" shows that there is a set of variables with "physical characteristics". The keywords for age of housing stock are "year" and "structure". Next, use {tidycensus} to explore the variables available:
```{r}
v17 <- load_variables(2022, 
                      "acs5",
                      cache = TRUE)


v17 |> 
  filter(str_detect(concept, "Year") & str_detect(concept, "Structure"))
```

Refine the search:
```{r}
v17 |> 
  filter(str_detect(label, "Estimate!!Total:!!Built"))
```

```{r}
housing_block_group <- get_acs(geography = "cbg",
                               state = "AZ",
                               variables = v17 |> 
                                 filter(str_detect(label, "Estimate!!Total:!!Built")) |>
                                 pull(name),
                               year = 2022,
                               geometry = TRUE)
```

Filter by urban Phoenix
```{r}
housing_block_group <- housing_block_group |>
  st_intersection(phoenix_urban)
```

The proportion of buildings built during certain periods could be a good indicator of post-war suburban development.

Save data objects:
```{r}
# Tracts
save(phoenix_tracts, 
     file = glue::glue(here::here(),
                       "/data-raw/phoenix_tracts.rda"),
     compress = "bzip2")
# Urban boundaries
save(phoenix_urban, 
     file = glue::glue(here::here(),
                       "/data-raw/phoenix_urban.rda"),
     compress = "bzip2")
# CBSA boundary
save(phoenix_cbsa, 
     file = glue::glue(here::here(),
                       "/data-raw/phoenix_cbsa.rda"),
     compress = "bzip2")

# h3 identifiers from which the coordinates can be retrieved for routing
save(phoenix_h3,
     file = glue::glue(here::here(),
                       "/data-raw/phoenix_h3.rda"),
     compress = "bzip2")

# h3 identifiers from which the coordinates can be retrieved for routing
save(housing_block_group,
     file = glue::glue(here::here(),
                       "/data-raw/housing_block_group.rda"),
     compress = "bzip2")
```



