---
title: "Removed Sections"
output: html_notebook
---

This notebook will contain sections removed from the city data notebooks found in `fortress-suburbia/data`. They are saved here in case they can be used in the future if/when deemed necessary.

# Previous Code for Removing Areas Disconnected from the Core Regions of the CBSAs:

## Phoenix, Arizona

Notice that there is a big urban area that is mostly outside of the Phoenix CBSA (Tucson). This will need to be removed.

```{r}
phoenix_urban <- phoenix_urban |>
  filter(!str_detect(NAME10, "Tucson"))
```

Verify:

```{r}
ggplot() +
  geom_sf(data = phoenix_cbsa,
          fill = "lightgray") +
  geom_sf(data = phoenix_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

## Houston, Texas

Notice that there is a small urban area that is south of the Houston CBSA (Lake Jackson), and small locations west of Houston (Prairie View, Pine Island, Hempstead). This will need to be removed.

```{r}
houston_urban <- houston_urban |>
  filter (!str_detect(NAME10, "Lake Jackson")) 

houston_urban <- houston_urban |>
  filter (!str_detect(NAME10, "Prairie View")) 

houston_urban <- houston_urban |>
  filter (!str_detect(NAME10, "Pine Island")) 

houston_urban <- houston_urban |>
  filter (!str_detect(NAME10, "Hempstead")) 

# Was unable to complete this within one filter line of code.
```

Verify:

```{r}
ggplot() +
  geom_sf(data = houston_cbsa,
          fill = "lightgray") +
  geom_sf(data = houston_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

## Atlanta, Georgia:

Notice that there are a few urban areas outside of Atlanta's boundaries. These will need to be removed.

```{r}
atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Calhoun")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Cartersville")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Jasper")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Nelson")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Yorkville")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Carrollton")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Bremen")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Grantville")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Shiloh")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Manchester")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Barnesville")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Jackson")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Monticello")) 

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Social Circle"))

atlanta_urban <- atlanta_urban |>
  filter (!str_detect(NAME10, "Madison")) 

# Was unable to complete this within one filter line of code.
```

Verify:

```{r}
ggplot() +
  geom_sf(data = atlanta_cbsa,
          fill = "lightgray") +
  geom_sf(data = atlanta_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```

## Nashville, Tennessee:

The small areas outside of the core Nashville area have to be removed. 

```{r}
nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Dickson")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Fairview")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Columbia")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Spring Hill")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Mount Pleasant")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Columbia")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Barfield")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Ashland City")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Pleasant View")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Lebanon")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Portland")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Murfreesboro")) #big blob

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Woodbury")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Columbia")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Carthage")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Lafayette")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "Springfield")) 

nashville_urban <- nashville_urban |>
  filter (!str_detect(NAME10, "White House"))  # ????????

# Was unable to complete this within one filter line of code.

# areas that connect by simply a line were also removed to ensure the h3 hashes worked more efficiently

# the location at ~36.4N, 86.8W (Greenbier & Ridegtop) were not possible to remove
```

Verify:

```{r}
ggplot() +
  geom_sf(data = nashville_cbsa,
          fill = "lightgray") +
  geom_sf(data = nashville_urban,
          color = "blue",
          fill = NA) +
  theme_minimal()
```